Early environmental conditions do not impact behavioural flexibility in an invasive and non-invasive lizard species

### 2024-05-30 ms.qmd
Updates

This study aimed to determine whether prenatal temperature and GCs impact behavioural flexibility in two closely related species, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). We expected incubation conditions to influence behavioural flexibility in both species; and we also predicted that _L. delicata_ would be more flexible than _L. guichenoti_ and less affected by the early environment. Nevertheless, we did not find significant differences between treatments or species when we compared their performance in a reversal task. Our results show that, on average, all the lizards from the four treatments and the two species learnt the reversal task - indicated by estimated slopes that are significantly different from zero - and that they learn, on average, at the same rate. Furthermore, these results were consistent when only those individuals who performed better in the previous associative task were included (see Supplementary material).

@@

More striking is the absence of an effect of incubation temperature on learning and behavioural flexibility because most studies have reported better performance in different associative learning and motor tasks when incubated at higher temperatures [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. Furthermore, those studies that reported cold-incubated lizards to learn faster [@dayananda_incubation_2017; @abayarathna_effects_2020] used incubation temperatures far beyond the natural range in their hot treatments. Our results suggest that behavioural flexibility is robust to temperature in both species of skinks, at least in our study. It could be arguable whether the differences in our incubation temperatures were extreme enough to elicit significant differences in learning performance. However, the mean difference between the temperatures used in our experiment  were similar to those in previous studies [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. In lizards, disparities in learning caused by the early thermal environment are associated to changes in neural density [@amiel_effects_2017], however, several studies have shown that the effects of temperature on brain structure and physiology are region dependent [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017]. It is possible that the temperatures used in our experiment did not affect the brain regions involved in coding reversal learning. Future studies should explore the effects of thermal environment on brain structure and function in these species to understand the mechanisms underlying the effects of temperature on learning. Furthermore, testing learning across different cognitive tasks could provide a more comprehensive understanding of the influence of temperature on cognition. In our analyses, sex was not included as a factor because the lizads employed were not considered sexually mature, and because previous studies did not find any differences on learning [@szabo2019sex], or on the effects of temperature on learning [@amiel_hotter_2012; @clark_colour_2014]. However, some studies evidence interactive effects of incubation temperature and sex on brain physiology in other species of reptiles [@coomber_independent_1997; @sakata_neural_2000]. In addition, the species employed by @amiel_hotter_2012, @amiel_egg_2014, and @clark_colour_2014, is known to sex-reverse - some individuals can have female chromosomes but male phenotypes - under cold temperatures [@dissanayake2021effects]; which was not accounted for in these experiments [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. As such, a potential temperature-sex interaction could be impacting our results, and future studies should explore the interactive effects of sex and incubation temperature on cognitive performance.

@@

@@

@@

Results for _L. delicata_ when the model included the age:
```{r, tbl-agedeli, results='asis', out.widht = "26cm"}
#| label: tbl-agedeli
#| tbl-cap: "Summary model_age_deli"
deli_age_res <- summary(model_age_deli)
# Extract fixed effects names and their corresponding estimates, CIs, etc.
fixed_effects <- deli_age_res$fixed %>%
  mutate_all(~ format_dec(., 2))
Predictors <- rownames(fixed_effects)
deli_age_tbl_df <- cbind(Predictors, fixed_effects)
# Convert the fixed effects data frame to a flextable
flextable(deli_age_tbl_df) %>%
  set_table_properties(., width = 1) %>%
  fontsize(size = 10, part = "all")
```

@@

Results for _L. guichenoti_ when the model included the age:
```{r, tbl-ageguich, results='asis'}
#| label: tbl-ageguich
#| tbl-cap: "Summary model_age_guich"
guich_age_res <- summary(model_age_guich)
# Extract fixed effects names and their corresponding estimates, CIs, etc.
fixed_effects <- guich_age_res$fixed %>%
  mutate_all(~ format_dec(., 2))
Predictors <- rownames(fixed_effects)
guich_age_tbl_df <- cbind(Predictors, fixed_effects)
# Convert the fixed effects data frame to a flextable
flextable(guich_age_tbl_df)%>%
  set_table_properties(., width = 1) %>%
  fontsize(size = 10, part = "all")
```

>>>>

This study aimed to determine whether prenatal temperature and GCs impact behavioural flexibility in two closely related species, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). We expected early environment to influence behavioural flexibility in both species. We also predicted that _L. delicata_ would be more flexible than _L. guichenoti_ and less affected by the early environment. Nevertheless, we did not find significant differences between treatments or species when we compared their performance in a reversal task. Our results show that, on average, all the lizards from the four treatments and the two species learnt the reversal task - indicated by estimated slopes that are significantly different from zero - and that they learn, on average, at the same rate. Furthermore, these results were consistent when only those individuals who performed better in the previous associative task were included (see Supplementary material).

@@

More striking is the absence of an effect of incubation temperature on behavioural flexibility because most studies have reported better performance in different associative learning and motor tasks when incubated at higher temperatures [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. Studies that have reported cold-incubated lizards to learning faster [@dayananda_incubation_2017; @abayarathna_effects_2020] used incubation temperatures far beyond the natural range in their hot treatments and it is unclear how such conditions relate to wild environments. Our results suggest that behavioural flexibility is robust to temperature in both species of skinks - at least in this study. It could be argued that the differences in our incubation temperatures were not extreme enough to elicit significant differences in learning performance. However, the mean difference between the temperatures used in our experiment  were similar to those in previous studies [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014], and are aligned with natural variation at the extremes of incubation conditions in the wild [@qualls2000post; @cheetham2011embryonic]. In lizards, impacts on learning from early thermal environments are associated with changes in neural density [@amiel_effects_2017]. However, several studies have shown that the effects of temperature on brain structure and physiology are region dependent [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017]. It is possible that the temperatures used in our experiment did not affect the brain regions involved in coding reversal learning. Future studies should explore the effects of thermal environment on brain structure and function in these species to understand the mechanisms underlying the effects of temperature on learning. Furthermore, testing learning across different cognitive tasks could provide a more comprehensive understanding of the influence of temperature on cognition. In our analyses, sex was not included as a factor because lizads were not considered sexually mature, and because previous studies have found weak evidence for sex-dependent learning in many cognitive domains [@szabo2019sex]. Indeed, the impacts of temperature on learning was not sex-depedent for the Eastern three-lined skink (_Bassiana dupeyerri_) [@amiel_hotter_2012; @clark_colour_2014]. However, some studies find evidence for interactive effects of incubation temperature and sex on brain physiology in other species of reptiles [@coomber_independent_1997; @sakata_neural_2000]. In addition, _B. dupeyerri_ are known to sex-reverse - some individuals can have female chromosomes but male phenotypes - under cold temperatures [@dissanayake2021effects]. Such effects were not easily accounted for in these experiments making it challenging to understand if sex or temperature were the major drivers of learning differences [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. Future studies should explore the possible interactive effects of sex and incubation temperature on cognitive performance. However, such studies will be challenging given the large sample sizes required to execute these studies rigorously.

@@

Table 1. Estimates of Reversal learning slope for all the different treatments per each task, species, and group. Mean shows the arithmetic means of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All pmcmc tested the hypothesis that the mean equals zero. In bold, those values that are significant (pmcmc <0.05)

@@

Table 2. Estimates of Reversal learning slope for all the different treatments per each task, species, and group. Here we included only those individuals who made the right choice in 8 out of the last 10 trials in the previous associative task. Mean shows the arithmetic means of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All pmcmc tested the hypothesis that the mean equals zero. In bold, those values that are significant (pmcmc <0.05)

@@

Table 3. Results for _L. delicata_ when the model included the age:
```{r, tbl-agedeli, results='asis', out.widht = "26cm"}
#| label: tbl-agedeli
#| tbl-cap: "Summary model_age_deli"
deli_age_res <- summary(model_age_deli)
# Extract fixed effects names and their corresponding estimates, CIs, etc.
fixed_effects <- deli_age_res$fixed %>%
  mutate_all(~ format_dec(., 2))
Predictors <- rownames(fixed_effects)
deli_age_tbl_df <- cbind(Predictors, fixed_effects)
# Convert the fixed effects data frame to a flextable
flextable(deli_age_tbl_df) %>%
  set_table_properties(., width = 1) %>%
  fontsize(size = 10, part = "all")
```

@@

Table 4. Results for _L. guichenoti_ when the model included the age:
```{r, tbl-ageguich, results='asis'}
#| label: tbl-ageguich
#| tbl-cap: "Summary model_age_guich"
guich_age_res <- summary(model_age_guich)
# Extract fixed effects names and their corresponding estimates, CIs, etc.
fixed_effects <- guich_age_res$fixed %>%
  mutate_all(~ format_dec(., 2))
Predictors <- rownames(fixed_effects)
guich_age_tbl_df <- cbind(Predictors, fixed_effects)
# Convert the fixed effects data frame to a flextable
flextable(guich_age_tbl_df)%>%
  set_table_properties(., width = 1) %>%
  fontsize(size = 10, part = "all")
```
